% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  10pt,
  american,
]{article}
\usepackage[]{Alegreya}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{multicol}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Table of Contents},
  pdfauthor={Yochai Gal},
  pdflang={en-US},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[papersize={5.7in,8.7in},margin=.5in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\else
  \usepackage[shorthands=off,main=american]{babel}
\fi

\title{Cairn Bestiary}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Compiled on July 14, 2022 by Yochai Gal | CC-BY-SA 4.0}
\date{}

\begin{document}
\maketitle
\begin{multicols}{2}
\let\oldtoc\tableofcontents
\renewcommand{\tableofcontents}{\oldtoc\newpage}
{
\setcounter{tocdepth}{3}
\tableofcontents
}
\end{multicols}
\widowpenalty=10000
\clubpenalty=10000

\hypertarget{acolyte}{%
\section{Acolyte}\label{acolyte}}

4 HP, 1 Armor, 14 WIL, mace (d6), Holy Symbol (Ward once per day)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Holy men and women in a quest for their deity.
\item Normally travel in groups of 4+.
\end{itemize}
\end{samepage}

\hypertarget{ankheg}{%
\section{Ankheg}\label{ankheg}}

7 HP, 2 Armor, 16 STR, 8 WIL, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge insectoids with multiple legs and shiny black eyes.
\item Stay burrowed just beneath the surfaces to ambush creatures walking above.
\item Squirts acid on distant enemies for d10 damage.
\end{itemize}
\end{samepage}

\hypertarget{ape-white}{%
\section{Ape, White}\label{ape-white}}

5 HP, 14 STR, claws (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Albino gorillas with nocturnal habits.
\item Defend their territory with threats and then violence.
\item Throw rocks at enemies for d6 damage.
\end{itemize}
\end{samepage}

\hypertarget{bandit}{%
\section{Bandit}\label{bandit}}

4 HP, 14 DEX, dagger (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Thieves who will value money over anything.
\item Use disguises, stealth, and trickery to surprise victims.
\end{itemize}
\end{samepage}

\hypertarget{banshee}{%
\section{Banshee}\label{banshee}}

8 HP, 3 Armor, 6 STR, 12 DEX, 15 WIL, ghostly touch (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Incorporeal feminine spirits that lingered after death and haunt the living.
\item Unharmed by cold, heat, or the elements.
\item Wail: Releases a piercing screech, anyone caught saves against \textbf{Critical Damage} regardless of current HP.
\end{itemize}
\end{samepage}

\hypertarget{basilisk}{%
\section{Basilisk}\label{basilisk}}

6 HP, 12 STR, 13 DEX, bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Long, magical, serpentine lizards that dwell in caverns.
\item Anyone who looks directly at it is slowed, having its attacks \emph{Impaired}.
\item \textbf{Critical Damage}: the victim loses d12 DEX, if reduced to 0 DEX, they are turned to pure stone.
\end{itemize}
\end{samepage}

\hypertarget{bat-vampire}{%
\section{Bat, Vampire}\label{bat-vampire}}

3 HP, 8 STR, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Big, flying mammals with nocturnal habits that nest in caves.
\item Unaffected by darkness or blinding effects, blinded by magical silence or constant loud noises.
\item Recover 1d4 HP after biting a victim that has blood.
\end{itemize}
\end{samepage}

\hypertarget{bear-black}{%
\section{Bear, Black}\label{bear-black}}

5 HP, 14 STR, claws (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 6’ tall bears who eat mostly berries and roots.
\item Only attack if cornered or to protect their young.
\item Sometimes raids camps for food.
\end{itemize}
\end{samepage}

\hypertarget{bear-cave}{%
\section{Bear, Cave}\label{bear-cave}}

8 HP, 17 STR, claws (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Ferocious 15’ tall carnivorous bears who live in caves.
\item Have poor eyesight, but track using smell.   Follow tracks of blood.
\item \textbf{Critical Damage}: Hug victims for d8 damage.
\end{itemize}
\end{samepage}

\hypertarget{bear-grizzly}{%
\section{Bear, Grizzly}\label{bear-grizzly}}

6 HP, 15 STR, claws (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Aggressive, 9’ tall reddish-brown furred bears who live in mountains.   Prefer to eat meat.
\item Frequently are found sleeping.
\end{itemize}
\end{samepage}

\hypertarget{bear-polar}{%
\section{Bear, Polar}\label{bear-polar}}

7 HP, 16 STR, claws (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 11’ tall white-furred bears who live in cold regions, eat mostly fish.
\item Excellent at swimming.
\item Can move effortlessly through snow and ice.
\end{itemize}
\end{samepage}

\hypertarget{beetle-fire}{%
\section{Beetle, Fire}\label{beetle-fire}}

2 HP, 4 STR, 12 DEX, 6 WIL, bite(d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 3’ long beetles commonly found underground.
\item Emits light through glands located in the abdomens and above the eyes.
\end{itemize}
\end{samepage}

\hypertarget{beetle-tiger}{%
\section{Beetle, Tiger}\label{beetle-tiger}}

3 HP, 6 STR, 12 DEX, 6 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 4’ long striped beetles with strong mandibles.
\item Eats other smaller creatures but sometimes hunts humans.
\end{itemize}
\end{samepage}

\hypertarget{berserker}{%
\section{Berserker}\label{berserker}}

4 HP, 1 Armor, 14 STR, battleaxe (d10, bulky)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Tribal fighters who fight recklessly, never taking prisoners.
\item Can enter in battle rage, enhancing all damage dealt and received.
\end{itemize}
\end{samepage}

\hypertarget{black-pudding}{%
\section{Black Pudding}\label{black-pudding}}

10 HP, 5 DEX, devouring touch (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, brainless blobs of black jelly driven by unstoppable hunger.
\item Can only be harmed by flames.
\item Can move in walls and ceilings and squeeze through small holes.
\end{itemize}
\end{samepage}

\hypertarget{blink-dog}{%
\section{Blink Dog}\label{blink-dog}}

3 HP, 14 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent, fey-like dogs that live in packs.
\item Teleport away in a blink after attacking.
\end{itemize}
\end{samepage}

\hypertarget{blood-elk}{%
\section{Blood Elk}\label{blood-elk}}

4 HP, 12 STR, 13 DEX, 5 WIL, horns (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Born from a greedy, selfish act.
\item Attacks and kills to eat meat, but gains no sustenance.
\item \textbf{Critical Damage}: gores its victims by ripping out their entrails.
\end{itemize}
\end{samepage}

\hypertarget{boar}{%
\section{Boar}\label{boar}}

3 HP, 12 STR, 6 WIL, tusks (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Omnivorous wild boars that dwell primarily in forests.
\item Not naturally aggressive, but can be dangerous if disturbed.
\end{itemize}
\end{samepage}

\hypertarget{boggart}{%
\section{Boggart}\label{boggart}}

3 HP, 4 STR, 17 DEX, 13 WIL

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item A wild, hairy trickster that takes pleasure in being a minor nuisance.
\item Prizes relics and shiny trinkets above all else but unwilling to trade for coin.
\item Boggarts have names that describe their true nature. Knowing their true name allows one to control a Boggart.
\end{itemize}
\end{samepage}

\hypertarget{brigand}{%
\section{Brigand}\label{brigand}}

4 HP, 1 Armor, shortsword (d6) or short bow (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Outlaws and mercenaries who raid settlements and ambush travelers.
\item Travel in large groups of at least one detachment.
\item For each detachment of regular brigands, there’s one leader who has chainmail (2 Armor) and a longsword (d8).
\end{itemize}
\end{samepage}

\hypertarget{buccaneer}{%
\section{Buccaneer}\label{buccaneer}}

4 HP, 14 DEX, 12 WIL, scimitar (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Outlaw sailors who raid coastal settlements and board other ships.
\item Travel in large ship crews, only fighting in clear advantage, generally numerical.
\end{itemize}
\end{samepage}

\hypertarget{bugbear}{%
\section{Bugbear}\label{bugbear}}

4 HP, 1 Armor, 14 DEX, club (d8, bulky)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, hairy, goblinoid-like creatures.
\item Use stealth to surprise and attack their victims.
\end{itemize}
\end{samepage}

\hypertarget{camel}{%
\section{Camel}\label{camel}}

3 HP, 15 STR, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Beasts of burden native to dry lands.
\item Might go weeks without water after being properly sated.
\item Moves through sand and broken ground without difficulty.
\end{itemize}
\end{samepage}

\hypertarget{cat-lion}{%
\section{Cat, Lion}\label{cat-lion}}

5 HP, 12 STR, 12 DEX, bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Hunts in groups called pride. Live in savannahs or other similar drylands.
\item Carnivorous. Prefer to eat bovines but might hunt even humans.
\item Always chase prey that flees.
\end{itemize}
\end{samepage}

\hypertarget{catoblepas}{%
\section{Catoblepas}\label{catoblepas}}

7 HP, 13 STR, tail(d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Monstrous creatures with a long, bony tail, a bloated body, short and stumpy legs, a long neck, and a boar-like head.   Dwell in swamps.
\item \textbf{Critical Damage}: the tail renders the victim stunned (d8 DEX damage).
\end{itemize}
\end{samepage}

\hypertarget{cat-panther}{%
\section{Cat, Panther}\label{cat-panther}}

4 HP, 14 DEX, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Dark furred cats that live on forests and plains.
\item Hunts medium or small-sized animals, using their high speed in their favor.
\end{itemize}
\end{samepage}

\hypertarget{cat-sabre-toothed-tiger}{%
\section{Cat, Sabre-Toothed Tiger}\label{cat-sabre-toothed-tiger}}

8 HP, STR 15, bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, aggressive cats with foot-long fangs.
\item Extremely rare, generally found in regions untouched by civilization.
\end{itemize}
\end{samepage}

\hypertarget{cat-tiger}{%
\section{Cat, Tiger}\label{cat-tiger}}

6 HP, 14 STR, 14 DEX, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, with striped fur.   Live in woodlands and colder regions.
\item Hunt alone, using their coloration to camouflage and surprise their victims.
\end{itemize}
\end{samepage}

\hypertarget{cave-locust}{%
\section{Cave Locust}\label{cave-locust}}

2 HP, 6 STR, 12 DEX, bite (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant, herbivorous crickets that dwell in caves.   Easily blend into stone due to their natural coloration.
\item Emits loud shrieks when threatened, possibly alerting other nearby creatures.
\item Immune to most types of poison.
\end{itemize}
\end{samepage}

\hypertarget{centaur}{%
\section{Centaur}\label{centaur}}

4 HP, 1 Armor, 12 STR, 12 DEX, 13 WIL, spear or short bow (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Creatures with the lower body and legs of a horse and the upper body and head of a human.
\item Live in small tribes hidden amidst dense forests with labyrinthic paths.
\end{itemize}
\end{samepage}

\hypertarget{centipede-giant}{%
\section{Centipede, Giant}\label{centipede-giant}}

1 HP, 6 STR, 6 WIL, bite (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 1’ long centipedes that live in damp, underground places.
\item \textbf{Critical Damage}: their deadly poison hinders the target incapacitated for up to 10 days (save STR once a day to recover).
\end{itemize}
\end{samepage}

\hypertarget{chimera}{%
\section{Chimera}\label{chimera}}

9 HP, 14 STR, bite (d8+d8, \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Three-headed flying monstrosities, hybrid of lion, goat, and dragon.
\item The draconic head breaths fire in a cone for d10 damage.
\end{itemize}
\end{samepage}

\hypertarget{cobblehounds}{%
\section{Cobblehounds}\label{cobblehounds}}

12 HP, 2 Armor, 14 STR, 1 DEX, 8 WIL, bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Immobile constructs typically used as guardians to great tombs or artifacts.
\item Unaffected by mundane persuasion techniques, but does love a good bone.
\end{itemize}
\end{samepage}

\hypertarget{cockatrice}{%
\section{Cockatrice}\label{cockatrice}}

5 HP, 8 STR, 12 DEX, beak (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Small, hybrid creatures resembling chickens with reptilian features.
\item \textbf{Critical Damage}: the victim loses d12 DEX, if reduced to 0 DEX, they are turned to pure stone.
\end{itemize}
\end{samepage}

\hypertarget{couatl}{%
\section{Couatl}\label{couatl}}

9 HP, 15 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Long, feathered serpents with colorful wings.   Live in hot jungles and are considered divine by native people.
\item Highly intelligent and magical, are capable of speaking multiple languages.
\item Shapechanger: may take the form of beasts and humanoids.
\end{itemize}
\end{samepage}

\hypertarget{crab-giant}{%
\section{Crab, Giant}\label{crab-giant}}

3 HP, 3 Armor, 11 STR, 4 DEX, pincers (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge crustaceans that live in coastal areas and mindlessly attack anything that comes near.
\item Their heavy carapace makes them notably sturdy but slow and unable to swim.
\end{itemize}
\end{samepage}

\hypertarget{crocodile-giant}{%
\section{Crocodile, Giant}\label{crocodile-giant}}

15 HP, 1 Armor, 17 STR, 8 DEX, bite (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 50’ long, extremely rare reptiles found on unexplored fluvial regions.
\item Attracted to the scent of blood or movement in the water.
\item Can easily destroy smaller ships with their crushing bite.
\end{itemize}
\end{samepage}

\hypertarget{crocodile-large}{%
\section{Crocodile, Large}\label{crocodile-large}}

6 HP, 1 Armor, 14 STR, 8 DEX, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 20’ long, may attack small boats and canoes.
\item Attracted to the scent of blood or heavy movement in the water.
\item Swim silently just under the surface to ambush victims.
\end{itemize}
\end{samepage}

\hypertarget{crypt-thing}{%
\section{Crypt Thing}\label{crypt-thing}}

12 HP, 3 Armor, 8 STR, 11 DEX, 14 WIL, ethereal claws (1d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item An animated skeleton clothed in billowing robes. Its eye sockets are hypnotically red.
\item Defenders of crypts and tombs, and will not attack if left undisturbed.
\item No armor against magical attacks.
\item Casts Teleport at will. Anyone in eyesight is teleported to a random room in the crypt.
\end{itemize}
\end{samepage}

\hypertarget{cyclops}{%
\section{Cyclops}\label{cyclops}}

13 HP, 18 STR, 8 DEX, 6 WIL, club (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant humanoids with a single large eye centered on the face.   Dwell in caves and herd small animals.
\item Slow-witted, can easily be fooled by other intelligent beings.
\end{itemize}
\end{samepage}

\hypertarget{deep-ones}{%
\section{Deep Ones}\label{deep-ones}}

4 HP, 11 STR, 8 DEX, 12 WIL, dagger (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Amphibious, fish-headed humanoids with webbed fingers.   Live in deep seas and are able to see in the darkness.
\item Highly resistant to magical effects.
\item Occasionally surface to trade with coastal settlements, frequently requesting union between one of them and a human.
\end{itemize}
\end{samepage}

\hypertarget{demon-knight}{%
\section{Demon Knight}\label{demon-knight}}

10 HP, 3 Armor, 16 STR, 8 DEX, 12 WIL, longsword (d8), Unholy Symbol
(Hatred when bathed in blood)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Once holy warriors who fell to the temptation of dark powers that they now serve in undeath.   Rides nightmarish steeds into battle.
\item Lesser undead creatures mindlessly follow their command.
\item Their evil aura frightens anyone who comes into nearby range.
\end{itemize}
\end{samepage}

\hypertarget{devil-swine}{%
\section{Devil Swine}\label{devil-swine}}

9 HP, 16 STR, 8 DEX, gore (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Corpulent humans who can change into massive swine. Lurk in isolated human settlements, hunting the people to eat their flesh.
\item Can only shape change at night, hunt in the dark, surprising victims.
\item \textbf{Critical Damage}: A human victim gets infected, becoming a lycanthrope of the same type after a couple of weeks.
\end{itemize}
\end{samepage}

\hypertarget{djinn}{%
\section{Djinn}\label{djinn}}

10 HP, 1 Armor, 15 DEX, fists (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Tall, cloudy humanoids from the plane of air.   Highly intelligent, use invisibility and illusions to trick other beings.
\item May transform into a whirlwind, sweeping everything on its path.
\item Capable of granting wishes, the result always is distorted based on the wording of the wish.
\end{itemize}
\end{samepage}

\hypertarget{dog-hunting}{%
\section{Dog, Hunting}\label{dog-hunting}}

2 HP, 12 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Bulky, domestic-only breeds with a ferocious nature.
\item Track by scent.   Once started, very difficult to put off the trail.
\item Only attack at their owner’s command.
\end{itemize}
\end{samepage}

\hypertarget{dog-war}{%
\section{Dog, War}\label{dog-war}}

3 HP, 1 Armor, 12 STR, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large breeds selected by their bulk and strength.
\item Not scared by noise or battle.
\end{itemize}
\end{samepage}

\hypertarget{dog-wild}{%
\section{Dog, Wild}\label{dog-wild}}

3 HP, 12 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Wild dogs that roam in forests in large packs.
\item A Wild dog automatically succeeds WIL saves while among its pack.
\end{itemize}
\end{samepage}

\hypertarget{dragon-black}{%
\section{Dragon, Black}\label{dragon-black}}

16 HP, 1 Armor, 13 STR, 18 DEX, 14 WIL, bite (d10), claws (d8+d8),
detachment

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item A giant amphibious reptile with glossy black scales. Dwells in swamps and other hostile flooded environments.
\item Extremely cruel and violent, makes use of the dark, difficult environment and its surprising agility to separate its victims and kill them one by one.
\item Any \textbf{Critical Damage} saves provoked by the dragon's bite automatically fail, due to its powerful acidic saliva.
\end{itemize}
\end{samepage}

\hypertarget{dragon-blue}{%
\section{Dragon, Blue}\label{dragon-blue}}

13 HP, 2 Armor, 16 STR, 17 WIL, bite (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Proud dragons that dwell in deserts and other drylands.
\item Extremely vain, often take victims as slaves to adore them.
\item Breathes lightning in a long line for d10 damage to all caught.
\end{itemize}
\end{samepage}

\hypertarget{dragon-brass}{%
\section{Dragon, Brass}\label{dragon-brass}}

11 HP, 2 Armor, 16 STR, 19 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Charismatic, good-aligned dragons who hide their hoards in dunes of sand.
\item Highly value conversations, frequently shapechanges into humanoids to interact with adventurers.
\item Breathes gas in a cone, targets affected fall asleep or flee in fear.
\end{itemize}
\end{samepage}

\hypertarget{dragon-bronze}{%
\section{Dragon, Bronze}\label{dragon-bronze}}

13 HP, 2 Armor, 16 STR, 14 DEX, 14 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}

\item Coastal dwelling, good dragons who eat mostly fish and aquatic plants.
\item Shapechange into small beasts to watch nearby wildlife and passing ships.
\item Breathes lightning on a line for d10 damage or repulsion gas in a cone, pushing victims away.
\end{itemize}
\end{samepage}

\hypertarget{dragon-copper}{%
\section{Dragon, Copper}\label{dragon-copper}}

12 HP, 2 Armor, 16 STR, 14 DEX, 18 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Good aligned dragons that dwell in dry, rocky regions.
\item Appreciate jokes and riddles, frequently prank passing travelers.
\item Breathes acid in a cone for d10 damage or slowing gas, halving the movement of all affected.
\end{itemize}
\end{samepage}

\hypertarget{dragon-gold}{%
\section{Dragon, Gold}\label{dragon-gold}}

15 HP, 3 Armor, 18 STR, 13 DEX, 16 WIL, bite (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Majestic, reserved dragons with dark golden scales. Can be found anywhere, but nest atop mountains, feeding of jewels and other minerals.
\item Master shapechangers, can take the form of any humanoid or animal.
\item Breath fire in a massive cone for d10 damage to all caught.
\end{itemize}
\end{samepage}

\hypertarget{dragon-green}{%
\section{Dragon, Green}\label{dragon-green}}

12 HP, 2 Armor, 15 STR, 18 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Treacherous, scheming forest-dwelling amphibious dragons.
\item Highly cunning, value the victims they tricked and their power of influence more than any hoard of gold.
\item Breathes a thick cloud of poisonous gas for d10 damage to all caught.
\end{itemize}
\end{samepage}

\hypertarget{dragon-red}{%
\section{Dragon, Red}\label{dragon-red}}

18 HP, 3 Armor, 18 STR, 12 DEX, 16 WIL, bite (d12), claws (d8+d8),
detachment

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item A gargantuan flaming creature, it's wingspan over a hundred feet. Arrogant and greedy, sees everything as its possession and everyone as its servant. Lairs within mountains with high volcanic activity.
\item Made of pure fire, is completely unharmed by any form of heat. Any cold vanishes within its presence.
\item Fire Breath: The dragon breathes a massive cone of pure flame dealing 12 damage (no roll) to all caught, the flame ignores and destroys any mundane armor. It needs a short rest before being able to do this again.
\end{itemize}
\end{samepage}

\hypertarget{dragon-silver}{%
\section{Dragon, Silver}\label{dragon-silver}}

13 HP, 3 Armor, 16 STR, 14 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Noble and selfless dragons, with shining scales rescinding of a knight's Armor. Dwell at the top of cold mountains.
\item Frequently deals with humans, going through great efforts to help those in need.
\item Breathe a line of cold dealing d10 damage and potentially paralyzing those caught.

\end{itemize}
\end{samepage}

\hypertarget{dragon-turtle}{%
\section{Dragon Turtle}\label{dragon-turtle}}

18 HP, 3 Armor, 18 STR, 6 DEX, 14 WIL, bite (d12+d12), detachment

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Gargantuan sea monsters with the shell of a turtle and draconic features. Lurk in deep oceans and can be mistaken for a small island when on the surface.
\item Attack ships, sinking them to add any treasure to their massive underwater hoards.
\item Breathes massive clouds of steam, clogging visions and damaging all inside for d10 damage.
\end{itemize}
\end{samepage}

\hypertarget{dragon-white}{%
\section{Dragon, White}\label{dragon-white}}

16 HP, 1 Armor, 16 STR, 16 DEX, 12 WIL, bite (d10), claws (d8+d8),
detachment

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item A massive winged creature of frost, it appears as if the winter itself coalesced into life. Wanders in the coldest regions, along with harsh snowstorms.
\item Extremely ferocious, is at the same time a clever hunter and a bestial predator. Tracks everything that lives, as if it was trying to satiate a endless hunger.
\item It's always surrounded by a storm of ice and cold wind, dealing d6 \emph{Blast} damage every round to everything nearby, in addition to its normal attacks.
\end{itemize}
\end{samepage}

\hypertarget{draugr}{%
\section{Draugr}\label{draugr}}

12 HP, 2 Armor, 15 STR, 9 DEX, 13 WIL, rusty broadsword (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Undead horror made of withered flesh. Rises from those killed in battle and left to rot.
\item \textbf{Critical Damage}: target is instantly  killed, only to rise later as a Thrall.
\end{itemize}
\end{samepage}

\hypertarget{drider}{%
\section{Drider}\label{drider}}

6 HP, 14 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Hybrid monstrosities resulting from the transformation ritual of the servants of the Drow spider deity.   Have the 8 legs of a giant spider and the upper body and head of the Drow.
\item Hunt on underworld caverns, eating the blood of paralyzed victims.   Frequently accompanied by giant spiders.
\item \textbf{Critical Damage}: The poison paralyzes the victim’s body (d6 DEX damage).
\end{itemize}
\end{samepage}

\hypertarget{driver-ant}{%
\section{Driver Ant}\label{driver-ant}}

3 HP, 12 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant, omnivorous black ants that build their colonies deep underground.
\item When hungry, they will consume anything in their path.
\item Once in melee combat, it will not stop attacking its victim until dead.
\end{itemize}
\end{samepage}

\hypertarget{drow}{%
\section{Drow}\label{drow}}

4 HP, 8 STR, 14 DEX, 12 WIL, shortsword (d6), a Spellbook (choose one:
Web, Spider Climb, or Night Sphere)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Elvish humanoids of dark skin and hair of silver and white.   Live deep in the underworld and feverishly worship strange deities.
\item Can see perfectly in the dark, but are blinded by bright light.
\item Wander in groups, frequently accompanied by giant spiders.
\end{itemize}
\end{samepage}

\hypertarget{dryad}{%
\section{Dryad}\label{dryad}}

4 HP, 14 WIL, unarmed (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Shy, peaceful tree spirits whose manifest in a beautiful female form.
\item Spiritually bound with a tree, may disappear by joining with the tree again. Can’t go too far away from it and dies if the tree is destroyed.
\item Charm strangers that come near. Charmed creatures mindlessly walk into the tree, disappearing forever if not rescued quickly.
\end{itemize}
\end{samepage}

\hypertarget{duergar}{%
\section{Duergar}\label{duergar}}

4 HP, 1 Armor, 14 STR, 8 WIL, Warhammer (d10, bulky)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Scrawny dwarven creatures with grey skin and ugly visages.   Dwell in underworld strongholds and are known for their avarice.
\item Can see perfectly in the dark, but are blinded by bright light.
\item Surprise victims using their ability to turn invisible for a short time.
\end{itemize}
\end{samepage}

\hypertarget{dwarf}{%
\section{Dwarf}\label{dwarf}}

4 HP, 2 Armor, 14 STR, hammer (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Short and bulky, bearded humanoids who dwell in mountains and underground.
\item Can see perfectly in low-light and partially in darkness.
\item Highly resistant to poison and disease.
\end{itemize}
\end{samepage}

\hypertarget{elemental-air}{%
\section{Elemental, Air}\label{elemental-air}}

16 HP, 11 STR, 15 DEX, 8 WIL, wind blow (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, living vortexes of whirling air.
\item Sweep away lighter creatures with their strong winds.
\item Harm aerial creatures by disrupting their flight.
\end{itemize}
\end{samepage}

\hypertarget{elemental-earth}{%
\section{Elemental, Earth}\label{elemental-earth}}

12 HP, 3 Armor, 16 STR, 8 DEX, fist (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge humanoid beings made of earth and stone.
\item Create small earth tremors with their natural strength.
\item Can meld into the earth and move through it as if swimming.
\end{itemize}
\end{samepage}

\hypertarget{elemental-fire}{%
\section{Elemental, Fire}\label{elemental-fire}}

16 HP, 12 DEX, 14 WIL, blow (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Living columns of pure flame that burn everything in their path.
\item Highly vulnerable to being touched by water or rain.
\item Can scatter their flames, freely changing its form.
\end{itemize}
\end{samepage}

\hypertarget{elemental-water}{%
\section{Elemental, Water}\label{elemental-water}}

12 HP, 2 Armor, 15 STR, blow (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge waves of flowing water.
\item Must stay near a body of water.
\item Envelop victims inside their forms to drown then.
\end{itemize}
\end{samepage}

\hypertarget{elephant}{%
\section{Elephant}\label{elephant}}

9 HP, 16 STR, 6 DEX, tusks (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large tusked animals that dwell near hot forests.   Found both alone and in herds.
\item With enough open ground for a clear run, it will charge its victims, enhancing damage.
\item Its tusks can be sold for high amounts of gold if removed whole.
\end{itemize}
\end{samepage}

\hypertarget{elf}{%
\section{Elf}\label{elf}}

4 HP, 1 Armor, 8 STR, 14 DEX, 14 WIL, shortswords (d6+d6) or longbow
(d8), a Spellbook (choose one: Charm or Detect Magic)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Slender, long-lived, fey humanoids with pointed ears. Live in harmony with nature, adapting to it instead of conquering.
\item Highly resistant to charming and mind-controlling effects.
\end{itemize}
\end{samepage}

\hypertarget{estrie}{%
\section{Estrie}\label{estrie}}

10 HP, 11 STR, 15 DEX, 14 WIL, draining hair (1d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Appears as a woman with long dark hair that extends all the way to her feet.
\item Can transform into owls, but only at night. Their shriek is ear-piercing.
\item Drinking blood gives them power, but they can obtain sustenance from eating bread and salt if taken from those they have wronged.
\item Binding their hair prevents their transformation. They can be killed, but their head must then be buried, the mouth stuffed with dirt.
\item \textbf{Critical Damage}: target falls unconscious for 1d4 days.

\end{itemize}
\end{samepage}

\hypertarget{ettin}{%
\section{Ettin}\label{ettin}}

10 HP, 16 STR, 6 WIL, club (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Two-headed giant kin of low intelligence and aggressive behavior.   Lay underground and only act in darkness.
\item One head is always vigilant, preventing being surprised.
\end{itemize}
\end{samepage}

\hypertarget{eye-of-terror}{%
\section{Eye of Terror}\label{eye-of-terror}}

12 HP, 12 STR, 8 DEX, 16 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Floating spheres with a large maw, one big central eye, and several small eyes on stalks scattered through the body.   Scheming and greedy, lair deep underground.
\item Cancels any magic effect in a nearby range facing its central eye.
\item The eyestalks cast one of those spells each round randomly: (1-Charm, 2-Phobia, 3-Telekinesis, 4-Sleep, 5-Shuffle, 6-Vision).
\end{itemize}
\end{samepage}

\hypertarget{flail-snail}{%
\section{Flail Snail}\label{flail-snail}}

4 HP, 2 Armor, 14 STR, 6 DEX, tentacles (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant snails with heavy, club-like tentacles and a colorful shell.   Dwell underground.
\item Its colorful shell can deflect magic, possibly annulling it or reflecting it to its caster.
\end{itemize}
\end{samepage}

\hypertarget{foxwoman}{%
\section{Foxwoman}\label{foxwoman}}

6 HP, 12 STR, 14 DEX, 11 WIL, teeth (d6), claws (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Appears as a 7-foot tall with a human woman with the head of a fox.
\item Transforms into a fox at will.
\end{itemize}
\end{samepage}

\hypertarget{frost-elf}{%
\section{Frost Elf}\label{frost-elf}}

14 HP, 1 Armor, 8 STR, 13 DEX, 14 WIL, icicle dagger (d6), a Spellbook
(Choose one: Sleep, Teleport, Detect Magic)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Beautiful, amoral, and long-lived.
\item Resistant to most forms of magic.
\end{itemize}
\end{samepage}

\hypertarget{gargoyle}{%
\section{Gargoyle}\label{gargoyle}}

6 HP, 3 Armor, 14 STR, 6 DEX, 12 WIL, claws (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Magical creatures that look like monstrous, horned, winged statues.
\item Guard places by standing still to pass like inanimate statues and attacking anything that comes near.
\item Cannot be charmed and put to sleep.
\end{itemize}
\end{samepage}

\hypertarget{gelatinous-cube}{%
\section{Gelatinous Cube}\label{gelatinous-cube}}

6 HP, 16 STR, 8 DEX, 4 WIL, touch (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large cubes of transparent jelly, dwell in dungeons absorbing items and living creatures.
\item May go undetected due to being transparent.
\item \textbf{Critical Damage}: Its touch paralyzes victims (d6 DEX damage).
\end{itemize}
\end{samepage}

\hypertarget{ghast}{%
\section{Ghast}\label{ghast}}

4 HP, 12 STR, WIL 6, claws (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Grotesque, humanoid undead that craves eating flesh. Found in packs of Ghouls.
\item Exhale a strong stench, nauseating nearby living creatures.
\item \textbf{Critical Damage}: The victim is paralyzed and loses 1d8 DEX. After paralyzing a victim, it moves to another target.
\end{itemize}
\end{samepage}

\hypertarget{ghost}{%
\section{Ghost}\label{ghost}}

10 HP, 3 Armor, 14 WIL, draining touch (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Incorporeal spirits of the restless dead.
\item Can possess victims that fail a WIL save, taking control of their bodies.
\item \textbf{Critical Damage}: Drains the victim’s life force (d6 extra STR damage).
\end{itemize}
\end{samepage}

\hypertarget{ghoul}{%
\section{Ghoul}\label{ghoul}}

6 HP, 11 STR, 14 DEX, 3 WIL, elongated claws (d6), jagged teeth (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Grey-skinned man-things wearing shredded clothes, hunkering over the ground.
\end{itemize}
\end{samepage}

\hypertarget{giant-catfish}{%
\section{Giant Catfish}\label{giant-catfish}}

8 HP, 14 STR, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Pale, 15’ long fish that lurk in the mud at the bottom of lakes and rivers.
\item Attack creatures that swim above their hiding spot.
\end{itemize}
\end{samepage}

\hypertarget{giant-cloud}{%
\section{Giant, Cloud}\label{giant-cloud}}

14 HP, 1 Armor, 16 STR, 12 DEX, 18 WIL, morningstar(d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 20’ tall humanoids with hair and skin in tones of grey and white. House in castles built atop the highest mountains or floating in cloud banks.
\item Cannot be surprised due to its keen sight and smell senses.
\item Can envelop itself in a wind aura, impairing missile attacks.
\end{itemize}
\end{samepage}

\hypertarget{giant-electric-eel}{%
\section{Giant Electric Eel}\label{giant-electric-eel}}

2 HP, 12 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 9’ long eels that live in warm, clear water.
\item Able to emit an electric shock, damaging all in the nearby water for d10.
\end{itemize}
\end{samepage}

\hypertarget{giant-fire}{%
\section{Giant, Fire}\label{giant-fire}}

13 HP, 3 Armor, 17 STR, 8 DEX, greatsword (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 16’ tall humanoids with red hair and charcoal skin. Wear heavy Armor made of brass, bronze or copper. Dwell in fortresses built near volcanoes.
\item Masters of craft and war, know how to forge the finest weaponry.
\item Immune to fire and heat.
\end{itemize}
\end{samepage}

\hypertarget{giant-frost}{%
\section{Giant, Frost}\label{giant-frost}}

13 HP, 2 Armor, 18 STR, 12 WIL, greataxe (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 18’ tall humanoids with blue-ish skin and pale hair. Wear furs and iron Armor.
\item Unharmed by cold and ice.
\item Excellent trackers, hunt accompanied by dire wolves.
\end{itemize}
\end{samepage}

\hypertarget{giant-hill}{%
\section{Giant, Hill}\label{giant-hill}}

12 HP, 15 STR, 6 WIL, great club (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 12’ tall, hairy and brutish humanoids. Dwell in hills and plains.
\item Low intelligence, relies only on its ferocity.
\item Frequently attack settlements for food.
\end{itemize}
\end{samepage}

\hypertarget{giant-pike}{%
\section{Giant Pike}\label{giant-pike}}

4 HP, 14 DEX, bite (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 10’ long, aggressive hunting fish. Dwell in deep lakes.
\item Use their great speed to surprise victims.
\end{itemize}
\end{samepage}

\hypertarget{giant-piranha}{%
\section{Giant Piranha}\label{giant-piranha}}

3 HP, 12 DEX, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 5’ long piranhas with colorful scales.   Live in rivers and attack anything in the water.
\item Dwell in large groups and overwhelm victims by swarming them.
\item Once it smells blood, it will not stop attacking.
\end{itemize}
\end{samepage}

\hypertarget{giant-rockfish}{%
\section{Giant Rockfish}\label{giant-rockfish}}

5 HP, 1 Armor, 12 STR, spine (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Spiny fish with rock-like scales. Live in saltwater. Normally passive, but attacks if disturbed.
\item Camouflages in rocks and coral formations.
\item \textbf{Critical Damage}: The spines release a highly lethal poison (d8 extra STR damage).
\end{itemize}
\end{samepage}

\hypertarget{giant-stone}{%
\section{Giant, Stone}\label{giant-stone}}

12 HP, 2 Armor, 16 STR, 14 DEX, stone club (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 14’ tall slender humanoids with stone-like gray skin.   Dwell in caverns or build primitive homes of stone in valleys.
\item Attack victims by throwing rocks at them from far away.
\item Can easily camouflage into stone.
\end{itemize}
\end{samepage}

\hypertarget{giant-storm}{%
\section{Giant, Storm}\label{giant-storm}}

18 HP, 2 Armor, 18 STR, 16 DEX, 18 WIL, greatsword (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 22’ tall humanoids with bronze-hued skin and light-colored hair. Live in mountain peaks and deep waters.
\item Unharmed by lightning, can summon storms at their will.
\item Can summon lightning bolts to attack nearby enemies for d12 damage.
\end{itemize}
\end{samepage}

\hypertarget{giant-sturgeon}{%
\section{Giant Sturgeon}\label{giant-sturgeon}}

10 HP, 2 Armor. 14 STR

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, 30’ long, aggressive armored fish.
\item Swallow its victims whole if able.
\end{itemize}
\end{samepage}

\hypertarget{giant-swordfish}{%
\section{Giant Swordfish}\label{giant-swordfish}}

3 HP, 12 DEX, sword (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Long saltwater fish with a bone sword on the snout. Mostly peaceful.
\item The bone sword might be used as a longsword(d8) if removed whole.
\end{itemize}
\end{samepage}

\hypertarget{gnoll}{%
\section{Gnoll}\label{gnoll}}

4 HP, 1 Armor, 12 STR, 8 WIL, spear or short bow (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Monstrous humanoid hyenas of low intelligence.   Legend says they were created in a wizard’s experiments.
\item Attack in large packs, intimidating its victims with numbers.
\item \textbf{Critical Damage}: the gnoll enters in a rampage after tasting blood, making another attack immediately.
\end{itemize}
\end{samepage}

\hypertarget{gnome}{%
\section{Gnome}\label{gnome}}

4 HP, 8 STR, 12 DEX, 14 WIL, crossbow (d8, bulky)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Short humanoids with long noses and pointed ears.   Live underground or deep in forests.
\item Highly intelligent tinkers.   Can easily be convinced by the promise of something valuable or uncommon.
\item Can see well in the darkness or low light.
\end{itemize}
\end{samepage}

\hypertarget{goblin}{%
\section{Goblin}\label{goblin}}

4 HP, 8 STR, 12 DEX, 8 WIL, dagger or short bow (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Small, grotesque humanoids with skin in earthly tones of green, brown, and grey.
\item Avoid combat, only attacking when in advantage, using hit-and-run tactics.
\item Sometimes are found using dire wolves as mounts.
\end{itemize}
\end{samepage}

\hypertarget{golem-amber}{%
\section{Golem, Amber}\label{golem-amber}}

10 HP, 3 Armor, 16 STR, 12 DEX, 14 WIL, claws (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Artificial constructs made in the shape of giant felines. Created by wizards in a slow and complex process.
\item Track without error, able to see invisible creatures nearby.
\item Immune to mundane attacks, gas, and mind-controlling effects.
\end{itemize}
\end{samepage}

\hypertarget{golem-bone}{%
\section{Golem, Bone}\label{golem-bone}}

8 HP, 2 Armor, 15 STR, 8 WIL, arms (d8, \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 6’ tall humanoid constructs made of human bone.   Have 4 arms spread around their torso.
\item Immune to mundane attacks, fire, cold, and electricity.
\end{itemize}
\end{samepage}

\hypertarget{golem-bronze}{%
\section{Golem, Bronze}\label{golem-bronze}}

18 HP, 3 Armor, 18 STR, 8 DEX, 11 WIL, fists (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Towering constructs made of pure bronze.   Reassemble fire giants, and emanate strong heat.
\item Immune to mundane attacks and fire.
\item When it suffers slashing or piercing damage to its STR, spurts a burning liquid, damaging the attacker for d8.
\end{itemize}
\end{samepage}

\hypertarget{golem-flesh}{%
\section{Golem, Flesh}\label{golem-flesh}}

9 HP, 1 Armor, 15 STR, 8 DEX, 8 WIL, fists (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 7’ tall humanoid abominations made of flesh stitched together.
\item Immune to mundane attacks, fire, and cold.
\item Absorb electricity damage received, healing HP by the same amount instead.
\item 
\end{itemize}
\end{samepage}

\hypertarget{golem-iron}{%
\section{Golem, Iron}\label{golem-iron}}

18 HP, 3 Armor, 18 STR, 8 WIL, sword (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 12’ humanoid statues forged with fire and magic.   Wield massive swords of iron.
\item Immune to mundane attacks, electricity, and cold.
\item Absorb electricity damage received, healing HP by the same amount instead.
\end{itemize}
\end{samepage}

\hypertarget{golem-original}{%
\section{Golem, Original}\label{golem-original}}

16 HP, 2 Armor, 15 STR, 6 DEX, 2 WIL, cobblestone fists (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item A hulking 9’ tall statue made of clay, mud, and prayer.
\item Mindlessly follows whatever commands given by its master, even after their death.
\item Immune to mundane blades and missile weapons. Damage from fire is \emph{Enhanced}.
\item If destroyed, the Golem eventually reforms and carries out its previous orders.
\end{itemize}
\end{samepage}

\hypertarget{golem-stone}{%
\section{Golem, Stone}\label{golem-stone}}

14 HP, 3 Armor, 17 STR, 8 DEX, fists (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 10’ tall animated statues carved from pure stone.
\item Immune to mundane attacks, fire, cold, and electricity.
\end{itemize}
\end{samepage}

\hypertarget{gorilla}{%
\section{Gorilla}\label{gorilla}}

4 HP, 14 STR, fists (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, muscular apes that dwell in hot jungles. Peaceful and Shy.
\item \textbf{Critical Damage}: Rends the victim for 1d4 extra STR damage.
\end{itemize}
\end{samepage}

\hypertarget{green-slime}{%
\section{Green Slime}\label{green-slime}}

3 HP, 16 STR, 8 DEX, 6 WIL, acidic touch (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large blobs of green slime that stick to walls and ceiling. Attack by dropping on top of their victims.
\item The acid corrodes both metal and wood along with the carrier, but cannot affect stone. Consumed flesh becomes more green slime.
\item Immune to all damage except fire. Once stuck on a victim, can only be removed if burned away.
\end{itemize}
\end{samepage}

\hypertarget{griffon}{%
\section{Griffon}\label{griffon}}

7 HP, 14 STR, 12 DEX, 9 WIL, beak (d8+d8) or claws (d6, \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Ravenous hybrid predators, with the head and wings of an eagle and the lower body of a lion.
\end{itemize}
\end{samepage}

\hypertarget{hag-black}{%
\section{Hag, Black}\label{hag-black}}

8 HP, 12 STR, 15 WIL, talons (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Hideous, 8' tall creatures that reassemble female humans of old age. Possess talons and teeth like iron and skin ranging from dark blue to black.
\item Dwell in dead forests and swamps, crave eating human flesh.
\item Can cloak their true appearance with an illusory form, and can see through any illusion.
\end{itemize}
\end{samepage}

\hypertarget{hag-sea}{%
\section{Hag, Sea}\label{hag-sea}}

3 HP, 12 STR, 14 WIL, knife (d6) or gaze (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Ghostly looking, green-skinned hags that dwell in seaweed forests and warm shallow waters. Crave eating human flesh.
\item Highly resistant to magic effects.
\item Hideous Gaze: Everyone facing the Hag must save WIL or drop to 0 HP.
\end{itemize}
\end{samepage}

\hypertarget{halfling}{%
\section{Halfling}\label{halfling}}

4 HP, 8 STR, 14 DEX, 14 WIL, shortswords (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Small humanoids with hairy feet live in small settlements called shires.
\item Highly resistant to fear effects.
\item Each shire is led by a stronger halfling called shire-riff, who has access to better equipment.
\end{itemize}
\end{samepage}

\hypertarget{harpy}{%
\section{Harpy}\label{harpy}}

5 HP, Armor 1, 12 STR, 12 DEX, 13 WIL, claws (d6+d6) or song (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Hideous monstrosities with the body of a giant eagle and a humanoid head.
\item Charming Song: Anyone who hears the song of a harpy must save WIL or be charmed, following the harpy and defending it from any threat. Killing the harpy breaks the charm.
\end{itemize}
\end{samepage}

\hypertarget{hawk-giant}{%
\section{Hawk, Giant}\label{hawk-giant}}

4 HP, 14 STR, 14 DEX, 8 WIL, talons or beak (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large birds of prey, the size of a big dog. Hunt animals of similar size, but might attack humans if hungry.
\item Surprise their victims by diving to attack from a great height.
\item Can carry away creatures up to their size.
\end{itemize}
\end{samepage}

\hypertarget{hawk-normal}{%
\section{Hawk, Normal}\label{hawk-normal}}

2 HP, 8 STR, 15 DEX, 8 WIL, talons or beak (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Small birds of prey. Hunt small rodents and reptiles, only attack larger creatures if they appear vulnerable.
\item Surprise their victims by diving to attack from a great height.
\item Can be trained as hunting companions.
\end{itemize}
\end{samepage}

\hypertarget{hellhound}{%
\section{Hellhound}\label{hellhound}}

5 HP, 1 Armor, 12 STR, 14 DEX, bite (d6) or fire breath (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Monstrous, intelligent dogs with the size of a pony. Dwell near volcanoes and sometimes accompany other fire-related creatures.
\item Immune to fire and heat. Can breathe fire, burning one target.
\item Don't rely on sight to find its foes, due to its keen sense of smell.
\end{itemize}
\end{samepage}

\hypertarget{hippogriff}{%
\section{Hippogriff}\label{hippogriff}}

3 HP, 12 STR, 14 DEX, claws (d6+d6) or bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Hybrid creatures with the head and front half of an eagle and the back half of a horse.
\item Can be trained to be used as mounts.
\item Hate pegasi, attacking then on sight.
\end{itemize}
\end{samepage}

\hypertarget{hippopotamus}{%
\section{Hippopotamus}\label{hippopotamus}}

8 HP, 1 Armor, 16 STR, 6 DEX, 8 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Massive, thick-skinned animals that live in tropical waters. Herbivore, but aggressively territorial.
\item Can hold their breath for long periods. Stay underwater and tip boats that come too close over.
\end{itemize}
\end{samepage}

\hypertarget{hobgoblin}{%
\section{Hobgoblin}\label{hobgoblin}}

4 HP, 2 Armor, 14 STR, 8 DEX, mace (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large and burly relatives of goblins. Dwell in underground fortresses, but frequently seek victims above ground.
\item Highly lawful and militarized culture, follow orders without question.
\item Trained to fight together, damage dealt is \emph{Enhanced} if an ally is also engaged with the same enemy.
\end{itemize}
\end{samepage}

\hypertarget{hooded-men}{%
\section{Hooded Men}\label{hooded-men}}

12 HP, 9 STR, 12 DEX, 14 WIL, leystaff (d8), a Spellbook (Choose one:
Charm, Hypnotize, Push/Pull, Shield)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item The Watchers of the Wood; a cult that derive their power from leylines, rune stones, and the like.
\item \textbf{Critical Damage}: leech a part of the victim's soul (1d4 WIL damage).
\end{itemize}
\end{samepage}

\hypertarget{hydra}{%
\section{Hydra}\label{hydra}}

9 HP, 1 Armor, 14 STR, 12 DEX, bite (d10, \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Draconian creatures with nine serpentine heads. Can attack a number of targets equal to the number of heads.
\item Each time it takes damage to STR, loses one head. Severed heads regrow after one turn, recovering 1 HP for head regrown.
\item Fire damage received is \emph{Enhanced}, and stops head regrowth for a turn.
\end{itemize}
\end{samepage}

\hypertarget{ifrit}{%
\section{Ifrit}\label{ifrit}}

10 HP, 15 STR, 14 WIL, flaming sword (d10, bulky)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Highly intelligent beings from the elemental plane of fire.  Manifests as huge men with terrifying faces and an aura of heat and smoke.
\item May transform into a pillar or wall of flames, burning everything it touches.
\item Might be summoned by powerful magic and ordered to complete tasks, but will subvert commands while following them to the letter.
\end{itemize}
\end{samepage}

\hypertarget{invisible-stalker}{%
\section{Invisible Stalker}\label{invisible-stalker}}

8 HP, 12 STR, 12 DEX, 15 WIL, blow (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Highly intelligent creatures summoned from other planes of existence by powerful wizards to perform tasks for them.
\item Can track any target without error. Using their invisibility to remain undetected.
\item If killed, is sent back to their original plane.
\end{itemize}
\end{samepage}

\hypertarget{jellyfish-giant}{%
\section{Jellyfish, Giant}\label{jellyfish-giant}}

4 HP, 12 STR, tentacles (d10,\emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, transparent sea creatures with a gelatinous body. Dwell in warm waters.
\item \textbf{Critical Damage}: The tentacles are covered in a poison that paralyzes the victim. (d8 DEX damage)
\item Killed victims are absorbed into the body and digested in a few minutes.
\end{itemize}
\end{samepage}

\hypertarget{killer-bee}{%
\section{Killer Bee}\label{killer-bee}}

4 HP, 6 STR, 12 DEX, 8 WIL, sting (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant bees that build hives underground. Hyper aggressive, attack anything that comes near.
\item Produce special honey that can heal d6 HP once per day if consumed.
\item \textbf{Critical Damage}: Leave their poisoned sting lodged in the target, inflicting d4 damage every round until removed. A killer bee dies soon after leaving its sting.
\end{itemize}
\end{samepage}

\hypertarget{kobold}{%
\section{Kobold}\label{kobold}}

4 HP, 8 STR, 13 DEX, dagger (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Small canine humanoids with hairless, scaly, red-brown skin. Dwell deep underground.
\item Can see perfectly in the dark, but are harmed by sunlight.
\item Ambush their victims in dark caves and tunnels.
\end{itemize}
\end{samepage}

\hypertarget{lamia}{%
\section{Lamia}\label{lamia}}

9 HP, 14 STR, 14 WIL, touch (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Centaur-like monsters with the head and upper body of a woman and the lower body of a beast. Prey on humanoids, drinking their blood and eating their flesh.
\item Can disguise their true form with illusions, appearing as a human woman.
\item \textbf{Critical Damage}: The touch saps the victim's wisdom (d6 WIL damage). A victim who reaches 0 WIL this way mindlessly obeys the lamia's command.
\end{itemize}
\end{samepage}

\hypertarget{lich}{%
\section{Lich}\label{lich}}

18 HP, 1 Armor, 8 STR, 12 DEX, 18 WIL, dagger (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Powerful wizards who refused death by turning themselves in soulless undead.
\item Able to cast virtually any spell
\item \textbf{Critical Damage}: Its touch paralyzes the victim (d8 DEX damage)
\end{itemize}
\end{samepage}

\hypertarget{lizard-giant-draco}{%
\section{Lizard, Giant~Draco}\label{lizard-giant-draco}}

4 HP, 12 STR, 14 DEX, bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 6' long carnivorous lizards with skin flaps between legs that allow gliding.
\item Dwell mostly on the surface, but sometimes can be found underground.
\end{itemize}
\end{samepage}

\hypertarget{lizard-giant-flame}{%
\section{Lizard, Giant Flame}\label{lizard-giant-flame}}

7 HP, 14 STR, 12 WIL, bite (d8+d8) or fire breath (d6+d6,\emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 30' long grey and red lizards that are sometimes mistaken by dragons. Lair underground but hunt on the surface. Sleep most of the day.
\item Unharmed by flames and heat.
\item Their eggs can be sold for a high price.
\end{itemize}
\end{samepage}

\hypertarget{lizard-giant-gecko}{%
\section{Lizard, Giant Gecko}\label{lizard-giant-gecko}}

3 HP, 12 DEX, 12 WIL, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 5' long, carnivorous, nocturn lizards. Blue and Orange colored.
\item Climb on sheer surfaces and attack by dropping on victims.
\end{itemize}
\end{samepage}

\hypertarget{lizard-giant-horned-chameleon}{%
\section{Lizard, Giant Horned
Chameleon}\label{lizard-giant-horned-chameleon}}

5 HP, 14 STR, 12 WIL, horn (d6) or tongue and bite (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 7' long lizards that camouflage by changing the color of their scales.
\item Use their camouflage to surprise victims.
\item Can attack further away targets by grabbing them with their elongated tongue. The victim must save STR or be pulled to the lizard's mouth and bitten.
\end{itemize}
\end{samepage}

\hypertarget{lizard-giant-monitor}{%
\section{Lizard, Giant Monitor}\label{lizard-giant-monitor}}

8 HP, 16 STR, 12 WIL, claws (d6+d6) or bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Gargantuan lizards with long teeth and claws. Dwell in warm climates.
\item Aggressive predators, surprise their victims by rushing at them and dragging them to their lair.
\item \textbf{Critical Damage}: The Victim is held into the mouth, automatically taking d6 damage in the following rounds.
\end{itemize}
\end{samepage}

\hypertarget{lizard-giant-subterranean}{%
\section{Lizard, Giant Subterranean}\label{lizard-giant-subterranean}}

6 HP, 15 STR, 12 WIL, bite (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 20' long, aggressive lizards that dwell only underground.
\item Can walk on walls and ceilings.
\item Clamps their mouth shut after biting a target, making it impossible to force open without killing the lizard.
\end{itemize}
\end{samepage}

\hypertarget{lizard-man}{%
\section{Lizard Man}\label{lizard-man}}

4 HP, 1 Armor, 14 STR, 12 DEX, bone spear (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Primitive, tribal, aquatic humanoids with reptilian heads and tails. Dwell in jungles nearby bodies of water.
\item Craft their Armors and weapons from the bones of their prey.
\item Carnivorous, hunt other humanoids, whose flesh they regard as a delicacy.
\end{itemize}
\end{samepage}

\hypertarget{locathah}{%
\section{Locathah}\label{locathah}}

4 HP, 12 STR, 12 WIL, spear (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Nomadic humanoid fish with pale scales and large eyes. Found in warm waters, lair in underwater caves.
\item Keep giant eels as pets, using them both as lair guardians and sometimes as mounts.
\end{itemize}
\end{samepage}

\hypertarget{manticore}{%
\section{Manticore}\label{manticore}}

6 HP, 15 DEX, 12 WIL, claws (d4+d4) or tail spike (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Monstrosities with a human face, a lion body, bat wings, and a spiked tail. Dwell in mountainous regions.
\item Prey on humans, following them and attacking with the spikes when they see an opening.
\item Attack with their tails by throwing the spikes like darts. The spikes regrow after a few days.
\end{itemize}
\end{samepage}

\hypertarget{mantid}{%
\section{Mantid}\label{mantid}}

6 HP, 1 Armor, 8 STR, 15 DEX, claws (d6+d6, \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent, 4-armed insectoid warriors who hunt and consume humanoids. Dwell in arid regions.
\item Can do huge leaps without needing any previous impulse.
\item Extremely skilled in dodging projectiles. Ranged damage received is \emph{Impaired}.
\end{itemize}
\end{samepage}

\hypertarget{mantis-giant}{%
\section{Mantis, Giant}\label{mantis-giant}}

10 HP, 1 Armor, 14 DEX, 12 WIL, claws (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 12' long insects with clawed forelimbs and slicing mandibles. Dwell in warm forests and jungles.
\item Hunt any prey, using their green coloration to camouflage in the foliage. Will not attack obviously stronger foes.
\item \textbf{Critical Damage}: Locks the victim in their claws, and bites their head off.
\end{itemize}
\end{samepage}

\hypertarget{mastodon}{%
\section{Mastodon}\label{mastodon}}

15 HP, 18 STR, 6 DEX, tusks (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge elephants with longer tusks. Found in icy regions.
\item With enough open ground for a clear run, it will charge its victims, enhancing damage.
\item Its tusks can be sold for high amounts of gold if removed whole.
\end{itemize}
\end{samepage}

\hypertarget{medusa}{%
\section{Medusa}\label{medusa}}

4 HP, 12 DEX, 16 WIL, snake bites (d6) or gaze (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Deadly magical creatures that look like women with snakes in place of hair.
\item Gaze: Anyone who looks directly at the medusa's face must save WIL or be turned to stone. Averting eyes from the face prevents this effect, but impairs any damage dealt.
\item \textbf{Critical Damage}: The snakes' poison paralyzes the victim (d6 DEX damage)
\end{itemize}
\end{samepage}

\hypertarget{merman}{%
\section{Merman}\label{merman}}

4 HP, 12 STR, 12 WIL, trident (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Aquatic humanoids with fishtails instead of legs. Live in coastal waters, farming seaweed, and hunting fish.
\item Train other sea creatures to guard their homes.
\end{itemize}
\end{samepage}

\hypertarget{mimic}{%
\section{Mimic}\label{mimic}}

9 HP, 2 Armor, 13 STR, 12 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Monstrous shape-shifters that take on the form of inanimate objects made of wood and stone. Found only underground.
\item Remain motionless (generally in the form of chests or doors), attacking anything that touches it.
\item Anything that touches the mimic is stuck in the glue that it exudes. Released only if the mimic is killed.
\end{itemize}
\end{samepage}

\hypertarget{mind-lasher}{%
\section{Mind Lasher}\label{mind-lasher}}

8 HP, 1 Armor, 8 STR, 12 DEX, 18 WIL, tentacle (d4), or mind
\emph{Blast} (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Humanoid creatures with purple skin and an octopus-like head. Dwell deep underground, plotting to slave humanity. Feed on the brains of other humanoids.
\item Mind \emph{Blast}: Emits a frequency that affects the brains of nearby creatures, all caught must save WIL or be paralyzed.
\item \textbf{Critical Damage}: The tentacle rips the victim's brain out of the skull, and the mind lasher eats it, gaining all of the victim's memories.
\end{itemize}
\end{samepage}

\hypertarget{minotaur}{%
\section{Minotaur}\label{minotaur}}

6 HP, 1 Armor, 15 STR, 8 WIL, axe (d10) or horns (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, muscular humanoids with the head of a bull. Found in mazes and labyrinths.
\item Attack other creatures on sight, pursuing until they are out of sight.
\end{itemize}
\end{samepage}

\hypertarget{mummy}{%
\section{Mummy}\label{mummy}}

5 HP, 12 STR, 8 DEX, 8 WIL, touch (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Undead humanoids wrapped in funerary bandages. Found in ruins of temples and tombs.
\item The first time anyone sees the mummy, it must save WIL or be paralyzed. The paralysis ends immediately if the mummy attacks or goes out of sight.
\item \textbf{Critical Damage}: The mummy inflicts a disease that rots open wounds, making them impossible to heal normally. Can only be removed by magic.
\end{itemize}
\end{samepage}

\hypertarget{nightmare}{%
\section{Nightmare}\label{nightmare}}

6 HP, 2 Armor, 14 STR, 14 DEX, 12 WIL, flaming hooves (d6+d6) or bite
(d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent demonic horses with burning-red eyes, smoldering nostrils, and flaming hooves.
\item Breathe a thick cloud of smoke, impairing damage dealt by anyone in melee with the nightmare.
\item Sometimes are used as steeds by powerful undead creatures.
\end{itemize}
\end{samepage}

\hypertarget{octopus-giant}{%
\section{Octopus, Giant}\label{octopus-giant}}

8 HP, 14 STR, 8 WIL, tentacles (d4+d4, \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, eight-armed saltwater creatures. Dwell nearby coastal settlements.
\item Can attack up to 8 nearby creatures.
\item When threatened, spills a thick cloud of ink and jets away.
\end{itemize}
\end{samepage}

\hypertarget{ogre}{%
\section{Ogre}\label{ogre}}

6 HP, 1 Armor, 15 STR, 8 DEX, 8 WIL, club (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 10' tall brute humanoids dressed in animal hides. Mostly dwell in caves.
\item Carry sacks full of gold and riches with them.
\end{itemize}
\end{samepage}

\hypertarget{orc}{%
\section{Orc}\label{orc}}

4 HP, 1 Armor, 14 STR, 8 WIL, axe (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Ugly humanoids with greyish skin, protuberant fangs, and powerful build. Animalistic and warmongering, hate other humanoids and delight in violence.
\item Can be convinced to join armies, under the promise of pillage.
\item Lair underground, sometimes along with ogres and trolls.
\end{itemize}
\end{samepage}

\hypertarget{owl-bear}{%
\section{Owl Bear}\label{owl-bear}}

7 HP, 15 STR, bite (d8) or claws (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, carnivorous bear-like creatures with the face of an owl. Found in deep dense forests.
\item Highly territorial, attack anything that comes to near.
\item \textbf{Critical Damage}: Hug victims for d8 extra damage.
\end{itemize}
\end{samepage}

\hypertarget{pegasus}{%
\section{Pegasus}\label{pegasus}}

3 HP, 12 STR, 14 DEX, 12 WIL. hooves (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent winged horses, wilful and proud. Seen as symbols of beauty and nobility.
\item Can be trained to serve as mounts for lawful characters.
\end{itemize}
\end{samepage}

\hypertarget{phoenix}{%
\section{Phoenix}\label{phoenix}}

18 HP, 3 Armor. 12 STR, 16 DEX, 14 WIL, talons (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant eagle-like birds made of pure flames, found in isolated regions and only attack in self-defense. Considered holy by some cultures.
\item Unharmed by fire and heat. Emits an aura of fire that does d6 damage to anyone that comes nearby.
\item If killed, disappears into a small sphere of flames. After a turn, it is reborn from its ashes and flees.
\end{itemize}
\end{samepage}

\hypertarget{pixie}{%
\section{Pixie}\label{pixie}}

12 HP, 6 STR, 15 DEX, 15 WIL, dagger (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Tiny humanoids with insectoid wings. Close relatives of fairies.
\item Naturally invisible, can reveal themselves if choose to.
\item Small wings only allow it to fly for short periods of time before needing to rest.
\end{itemize}
\end{samepage}

\hypertarget{pseudo-dragon}{%
\section{Pseudo-Dragon}\label{pseudo-dragon}}

3 HP, 8 STR, 14 DEX, 12 WIL, sting (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent, small winged draconic creatures with a sting-tipped tail and reddish scales. Found in forests and caverns.
\item Capable of communicating telepathically with nearby creatures.
\item \textbf{Critical Damage}: The sting's poison leaves the victim in a comatose state for 1d6 days.
\end{itemize}
\end{samepage}

\hypertarget{pteranodon}{%
\section{Pteranodon}\label{pteranodon}}

5 HP, 14 STR, 12 DEX, bite (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Massive winged reptiles that live in warm climates, hunting creatures on the ground by gliding on air currents.
\item Highly aggressive, often prey on humanoids.
\end{itemize}
\end{samepage}

\hypertarget{purple-worm}{%
\section{Purple Worm}\label{purple-worm}}

18 HP, 1 Armor, 18 STR, 8 DEX, 6 WIL, bite (d10+d10) or poison sting
(d8), Detachment

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Gargantuan worms with bodies more than 100' long and 10' thick. Burrow in tunnels underground and surface to eat other creatures. Found in deserts and other drylands.
\item Victims bitten by the worm must save STR or be swallowed whole. Anything inside the worm receives d12 acid damage each round.
\item \textbf{Critical Damage}: The sting inflicts a highly lethal poison, dealing d12 extra STR damage.
\end{itemize}
\end{samepage}

\hypertarget{rakshasa}{%
\section{Rakshasa}\label{rakshasa}}

7 HP, 12 DEX, 15 WIL, claws (d6+d6), 3 random spellbooks

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Cunning spirits that take the fort of great humanoid cats. Feed on human flesh, deceiving their victims with illusions and then attacking.
\item Can read the mind of other creatures and then cloak their true form with an illusory form that the creature deems friendly.
\item A blessed weapon instantly kills a rakshasa.
\end{itemize}
\end{samepage}

\hypertarget{red-cap}{%
\section{Red Cap}\label{red-cap}}

4 HP, 12 DEX, 8 WIL, two sickles (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Borne from blood left to rot in the fae. Hats colored by blood; they even bathe in blood.
\item Attack travelers, but will ignore those with little to live for, as there is no pleasure in it.
\item \textbf{Critical Damage}: eviscerate the target, showering in blood, restoring any lost STR and HP.
\end{itemize}
\end{samepage}

\hypertarget{remorhaz}{%
\section{Remorhaz}\label{remorhaz}}

11 HP, 1 Armor, 14 STR, 12 DEX, 8 WIL, bite (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Wormy-like creatures with multiple legs and eyes, long bodies, and strong, crushing maws. Dwell in frozen lands, sustained by their burning internal heat.
\item If their nests are found, eggs can be sold for high prices.
\item \textbf{Critical Damage}: The victim is swallowed whole, and instantly killed by the internal heat.
\end{itemize}
\end{samepage}

\hypertarget{rhinoceros}{%
\section{Rhinoceros}\label{rhinoceros}}

6 HP, 2 Armor, 16 STR, 8 DEX, horn (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge, armored, herbivorous mammals. Highly territorial, dwell in warm climates.
\item With enough open ground for a clear run, it will charge its victims, enhancing damage.
\end{itemize}
\end{samepage}

\hypertarget{roc-giant}{%
\section{Roc, Giant}\label{roc-giant}}

18 HP, 18 STR, 14 DEX, 12 WIL, claws (d8+d8, \emph{Blast}) or bite
(d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Gargantuan birds of prey that nest atop the highest peaks and attack anything that approaches their nests. Considered myths by most people.
\item Surprise victims by swooping down from above.
\item Can be domesticated if captured when young.
\end{itemize}
\end{samepage}

\hypertarget{root-goblin}{%
\section{Root Goblin}\label{root-goblin}}

4 HP, 8 STR, 14 DEX, 8 WIL, spear (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Avoid combat unless they have the advantage (such as greater numbers).
\item Guard their stolen goods to the death.
\item Prize Spellbooks; willing to trade.
\end{itemize}
\end{samepage}

\hypertarget{root-witch}{%
\section{Root Witch}\label{root-witch}}

8 HP, 9 STR, 16 DEX, 14 WIL, tuber-fingers (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Excellent tunneler that hides underground near water. Wiggles tiny fronds into the air as lures.
\item Above-ground, appears vaguely humanoid with the face of a rotted tree stump.
\item Will exchange rare minerals for a “fresh corpse” of any kind.
\end{itemize}
\end{samepage}

\hypertarget{rust-monster}{%
\section{Rust Monster}\label{rust-monster}}

5 HP, 12 DEX, 12 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Armadillo-like creatures with long tails and two long antennae. Feed on rusted metal.
\item Any metal that touches the antennae instantly becomes rusted. Magical items are partially resistant to this effect.
\item Can smell metal at long distances.
\end{itemize}
\end{samepage}

\hypertarget{sahuagin}{%
\section{Sahuagin}\label{sahuagin}}

3 HP, 12 STR, 12 DEX, trident (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Aquatic humanoids with black-green scales and deep black eyes, known in some places as "sea devils". Dwell in warm salt waters and raid coastal villages during the night.
\item Can breathe outside of water for up to 4 hours.
\item Command sharks to guard their lairs.
\end{itemize}
\end{samepage}

\hypertarget{salamander-flame}{%
\section{Salamander, Flame}\label{salamander-flame}}

8 HP, 12 DEX, 14 WIL, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 12' long lizard-like creatures with scales of bright orange, yellow and red. Native to the plane of fire, but also dwell near volcanoes. 
\item Immune to fire and heat.
\item Emits an aura of heat, damaging any nearby creature for d6 damage.
\end{itemize}
\end{samepage}

\hypertarget{satyr}{%
\section{Satyr}\label{satyr}}

5 HP, 12 DEX, 16 WIL, horns (d6) or pipes (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Magical humanoids with the legs and horns of a goat. Love to dance and drink. Found in deep forests.
\item Highly resistant to magic effects.
\item Pipes: Anyone who hears the music from a satyr's pipes must save WIL or fall asleep, be charmed or frightened. The Satyr chooses which effect when playing the pipes.
\end{itemize}
\end{samepage}

\hypertarget{scorpion-giant}{%
\section{Scorpion, Giant}\label{scorpion-giant}}

4 HP, 2 Armor, 12 STR, 12 DEX, 8 WIL, claws (d6+d6) or sting (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Huge arachnids, the size of a horse, with pincers and poisonous stingers. Found in drylands and caverns. Highly aggressive, normally attack on sight.
\item Immobilizes its victims with the claws, and then attack with the sting.
\item \textbf{Critical Damage}: The sting's poison instantly kills the target.
\end{itemize}
\end{samepage}

\hypertarget{shadow}{%
\section{Shadow}\label{shadow}}

4 HP, 8 STR, 14 WIL, touch (d6, ignores armor)~

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Incorporeal monsters that look like living shadows.
\item Unharmed by mundane attacks, sleep or mind control.
\item \textbf{Critical Damage}: The victim loses another d4 STR, if reduced to 0 STR, they become a shadow.
\end{itemize}
\end{samepage}

\hypertarget{shambling-mound}{%
\section{Shambling Mound}\label{shambling-mound}}

9 HP, 15 STR, 6 DEX, 8 WIL, fists (d8+d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Plant monsters with the rough shape of a 9' tall humanoid made of slimy vegetation. Dwell in dark swamps and damp undergrounds.
\item Damage received from weapons is \emph{Impaired}.
\item Entagles its victims to suffocate them in the roots and tendrils of its body.
\end{itemize}
\end{samepage}

\hypertarget{shark-great-white}{%
\section{Shark, Great White}\label{shark-great-white}}

8 HP. 14 STR, 14 DEX, 6 WIL, bite (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 30' long aggressive fish of a grey coloration. Dwell deep salt water, and sometimes attack smaller boats.
\item Can detect the smell of blood from great distances.
\item Once moves to attack, doesn't stop until death, succeeding all morale rolls.
\end{itemize}
\end{samepage}

\hypertarget{skeleton}{%
\section{Skeleton}\label{skeleton}}

5 HP, 1 Armor, 13 DEX, rusty sword (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item If a skeleton is killed and its bones are not scattered, it reforms.
\end{itemize}
\end{samepage}

\hypertarget{snake-giant-rattler}{%
\section{Snake, Giant Rattler}\label{snake-giant-rattler}}

4 HP, 14 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 10' long snakes with brown and white scales and a rattle of hardened scales on the tip of their tail.
\item Can attack two targets in the same round, due to its great speed.
\item \textbf{Critical Damage}: The poison kills the target in 2d6 turns if not treated.
\end{itemize}
\end{samepage}

\hypertarget{snake-person}{%
\section{Snake Person}\label{snake-person}}

6 HP, 1 Armor, 12 DEX, 14 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Creatures with the body, tail, and head of a giant snake and scaled humanoid torso and arms. Capture other humanoids to be eaten or enslaved by their cruel masters. Dwell in hot jungles.
\item Highly resistant to magic effects.
\item Supposedly can breed with humans, creating hybrids that have the appearance of humans with reptilian eyes and forked tongues.
\end{itemize}
\end{samepage}

\hypertarget{snake-pit-viper}{%
\section{Snake, Pit Viper}\label{snake-pit-viper}}

3 HP, 12 DEX, bite (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 5' long snakes with greyish scales, found in caves and other dark places.
\item Does not rely on vision, instead, senses its victims by their body heat.
\item \textbf{Critical Damage}: The poison kills the target in 2d4 turns if not treated.
\end{itemize}
\end{samepage}

\hypertarget{snake-rock-python}{%
\section{Snake, Rock Python}\label{snake-rock-python}}

5 HP, 14 STR, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 20' long, thick snakes with brown and yellow scales.
\item After a biting attack, attempts to constrict the victim with its body. The victim must save STR or be restrained, taking d8 damage every turn. 
\end{itemize}
\end{samepage}

\hypertarget{snake-spitting-cobra}{%
\section{Snake, Spitting Cobra}\label{snake-spitting-cobra}}

3 HP, 14 DEX, bite (d6) or acid spit (d4)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 3' long snake with grey and white scales. Prefer to attack from distance with their acidic spit.
\item \textbf{Critical Damage}: The poison kills the target in 2d8 turns if not treated.
\end{itemize}
\end{samepage}

\hypertarget{sphinx}{%
\section{Sphinx}\label{sphinx}}

12 HP, 12 STR, 15 WIL, claws (d6+d6) or roar (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large monsters with bird wings, the body of a lion, and a human face. Collect puzzles, riddles, and obscure knowledge. May pose riddles to those they meet, attacking and consuming any who cannot give the correct answer.
\item Roar: Anyone who hears the sphinx's roar must save WIL or flee in fear.
\item Can innately cast Comprehend, Detect Magic, and Telekinesis. Once per day each.
\end{itemize}
\end{samepage}

\hypertarget{spider-giant-aquatic}{%
\section{Spider, Giant Aquatic}\label{spider-giant-aquatic}}

4 HP, 12 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 4’ long spiders that dwell in large bodies of freshwater. Build air-filled nests in aquatic vegetation.
\item Lurk just below the water's surface to surprise its victims.
\item \textbf{Critical Damage}: The poison paralyzes the victim (d8 DEX damage), the spider drags paralyzed victims to its lair to eat them.
\end{itemize}
\end{samepage}

\hypertarget{spider-giant-aranea}{%
\section{Spider, Giant Aranea}\label{spider-giant-aranea}}

3 HP, 12 DEX, 14 WIL, bite (d6), two random spellbooks

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 6’ long, intelligent spiders of greenish-black coloration. Have an odd lump on the back, housing its large brain. Two front legs have digits, allowing an aranea to grasp tools. Dwell in web-filled lairs where they conduct magical research.
\item  Scared of fire, avoid using spells of such effect.
\item Creatures caught in webs are entangled and unable to move. Breaking free requires an STR save.
\end{itemize}
\end{samepage}

\hypertarget{spider-giant-black-widow}{%
\section{Spider, Giant Black Widow}\label{spider-giant-black-widow}}

3 HP, 14 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 6’ long, black spiders with a red hourglass pattern on their abdomens. Dwell in web-filled lairs and sometimes prey on humans.
\item Creatures caught in webs are entangled and unable to move. Breaking free requires an STR save.
\item \textbf{Critical Damage}: The poison kills the target in 2d4 turns if not treated.
\end{itemize}
\end{samepage}

\hypertarget{spider-giant-phase}{%
\section{Spider, Giant Phase}\label{spider-giant-phase}}

5 HP, 14 DEX, 12 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 8’ long, black spiders that have the ability to become intangible. Dwell in web-filled lairs and sometimes prey on humans.
\item Can shift out of phase with surroundings, becoming intangible. Attack by briefly phasing back in.
\item \textbf{Critical Damage}: The poison kills the target in 2d6 turns if not treated.
\end{itemize}
\end{samepage}

\hypertarget{svirneblin}{%
\section{Svirneblin}\label{svirneblin}}

4 HP, 1 Armor, 8 STR, 12 DEX, 14 WIL, adze (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Short, thickset demihumans with long noses and gnarled, hairless, grey skin. Also known as deep gnomes. Dwell deep in the Underworld in labyrinthine lairs riddled with cunning secret passages and chambers.
\item Can see perfectly in the dark, but are blinded by bright light.
\item Love machinery, mining, and gems.
\end{itemize}
\end{samepage}

\hypertarget{tarrasque}{%
\section{Tarrasque}\label{tarrasque}}

18 HP, 3 Armor, 18 STR, 8 WIL, claws (d12+d12), Detachment

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Legendary, gargantuan (50’ long) reptilian monstrosities with plated skin, dreaded throughout all lands. Spend most of their time sleeping in their underground lairs. Emerge periodically and ravage vast swaths of land, consuming all plants and animals in the path. Possibly the most feared monster of all.
\item Regenerates 1 HP every turn. A Tarrasque always regenerates, even if its body is reduced to pieces. The way of permanently killing it is unknown.
\item \textbf{Critical Damage}: The razor-sharp claws sever a part of the victim's body.
\end{itemize}
\end{samepage}

\hypertarget{titan}{%
\section{Titan}\label{titan}}

18 HP, 2 Armor, 16 STR, 12 DEX, 18 WIL, sword (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 20’ tall humanoids of radiant beauty and athletic build. Beneficent of character, but prone to megalomania. Natives of higher planes, but sometimes have strongholds in the mortal world.
\item Can become invisible and levitate at will. Know a large number of spells.
\item Frequently command storm giants as their guards.
\end{itemize}
\end{samepage}

\hypertarget{treant}{%
\section{Treant}\label{treant}}

8 HP, 1 Armor, 15 STR, 8 DEX, fists (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Giant humanoids made of tree bark and foliage. Dwell in forests. Are concerned only with protecting the plants of their home.
\item Hate fire, and attack anyone who wields it.
\item Surprise victims by standing still and passing for a normal tree.
\end{itemize}
\end{samepage}

\hypertarget{triton}{%
\section{Triton}\label{triton}}

4 HP, 1 Armor, 12 STR, 12 DEX, 12 WIL, trident (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Aquatic humanoids with silvery skin, blue-green hair, and scaled legs ending in fins. Dwell in warm waters both in shallows and at great depths.
\item Frequently use seahorses and hippocamps as mounts.
\item Lair in beautiful castles sculpted from sea rocks and corals, guarded by other sea creatures.
\end{itemize}
\end{samepage}

\hypertarget{troglodyte}{%
\section{Troglodyte}\label{troglodyte}}

4 HP, 12 DEX, claws (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent, reptilian humanoids with agile hands, long legs, short tails, and spiky combs on their heads and arms.
\item Can change color to match their surroundings. Lurk by rock walls and await victims.
\item Emits an oil with a nauseating stench from its pores.
\end{itemize}
\end{samepage}

\hypertarget{troll}{%
\section{Troll}\label{troll}}

6 HP, 1 Armor, 14 STR, 12 DEX, talons (d6+d6) or bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Wicked, 8’ tall humanoids with rubbery bodies. Consume the flesh of other humanoids. Dwell underground, in the barren  wilderness, and in the ruined homes of former victims.
\item Fire and acid damage received is \emph{Enhanced}, and stops a troll's regeneration. When attacked with either of those, the troll fails all morale checks.
\item Regains 3 HP per round, and even severed limbs are reattached. If killed, will regenerate and fight again in 2d6 minutes.
\end{itemize}
\end{samepage}

\hypertarget{turtle-giant-sea}{%
\section{Turtle, Giant Sea}\label{turtle-giant-sea}}

15 HP, 2 Armor, 15 STR, 6 DEX, 12 WIL, bite (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 20’ long, peaceful marine reptiles that drift in warm oceans, eating sea plants. Dangerous if threatened.
\item Tip over small boats that come into their territory.

\end{itemize}
\end{samepage}

\hypertarget{turtle-giant-snapping}{%
\section{Turtle, Giant Snapping~}\label{turtle-giant-snapping}}

10 HP, 14 STR, 8 DEX, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 10’ long, aggressive predatory turtles that lurk in the shallows of lakes and large rivers.
\item Hide motionless underwater to surprise its victims.
\item Can reach to bite further targets due to its long neck.
\end{itemize}
\end{samepage}

\hypertarget{tyrannosaurus-rex}{%
\section{Tyrannosaurus Rex}\label{tyrannosaurus-rex}}

18 HP, 1 Armor, 18 STR, 12 DEX, bite (d10+d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Massive, over 20’ tall, two-legged, predatory dinosaurs with huge jaws. Hunt human-sized or larger prey. Dwell in Lost World regions.
\item Always attack the largest, most threatening foe first.
\end{itemize}
\end{samepage}

\hypertarget{unicorn}{%
\section{Unicorn}\label{unicorn}}

4 HP, 12 DEX, 14 WIL, horn (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Magical horses with a single longhorn. Unicorns are timid, but proud and wilful. Seen as divine beings by many.
\item Can teleport to a visible distance once a day.
\end{itemize}
\end{samepage}

\hypertarget{vampire}{%
\section{Vampire}\label{vampire}}

12 HP, 1 Armor, 12 STR, 12 DEX, 16 WIL, bite (d10)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Charming, undead creatures that drink the blood of mortals. Act at night and sleep in a coffin during the day. Can shape change to the form of a giant bat.
\item Regeneration: A damaged vampire gains 3 HP at the start of each round, as long as it is alive. If killed, becomes a cloud of gas and retreats to its coffin, reforming at the next nightfall. Can only be killed if exposed to sunlight or if the coffin is destroyed.
\item \textbf{Critical Damage}: The Vampire drains the victim's essence (d12 WIL damage). At 0 WIL, the victim is killed and becomes a vampire.
\end{itemize}
\end{samepage}

\hypertarget{warp-beast}{%
\section{Warp Beast}\label{warp-beast}}

6 HP, 14 DEX, 12 WILL, tentacles (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, black, six-legged monsters. Look similar to panthers with a tentacle growing from each front shoulder. Tentacles are ridged with razor-sharp edges.
\item Displace to a short distance to avoid attacks, impairing the damage received.
\item Highly resistant to magical effects.
\end{itemize}
\end{samepage}

\hypertarget{weasel-giant}{%
\section{Weasel, Giant}\label{weasel-giant}}

4 HP, 12 DEX, 12 WIL, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item 9’ long, vicious, predatory mammals with rich fur of brown, gold, or white. Dwell in subterranean tunnels, hunting alone or in small groups.
\item After attacking, locks onto the victim and drains its blood for 1d8 automatic damage each round. Detaches only if the weasel or the victim dies.
\item Track by scent, preferring wounded victims.
\end{itemize}
\end{samepage}

\hypertarget{werebear}{%
\section{Werebear}\label{werebear}}

6 HP, 14 STR, 8 DEX, claws (d6+d6) or bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Highly intelligent, even in bear form. Frequently dwell among normal bears.
\item Can be reasoned with, if approached peacefully.
\item \textbf{Critical Damage}: A human victim gets infected, becoming a lycanthrope of the same type after a couple of weeks.
\end{itemize}
\end{samepage}

\hypertarget{wereboar}{%
\section{Wereboar}\label{wereboar}}

4 HP, 13 STR, 12 DEX, 8 WIL, tusks (d6+d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Highly aggressive, even in human form. Fight until death.
\item While in human form, can enter in battle rage, enhancing all damage dealt and received.
\item \textbf{Critical Damage}: A human victim gets infected, becoming a lycanthrope of the same type after a couple of weeks.
\end{itemize}
\end{samepage}

\hypertarget{wererat}{%
\section{Wererat}\label{wererat}}

3 HP, 8 STR, 14 DEX, 12 WIL, bite (d4+d4) or weapon (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Intelligent, humanoid rats that can change into normal humans.
\item Can speak and wield weapons normally while in rat form.
\item \textbf{Critical Damage}: A human victim gets infected, becoming a lycanthrope of the same type after a couple of weeks.
\end{itemize}
\end{samepage}

\hypertarget{weretiger}{%
\section{Weretiger}\label{weretiger}}

5 HP, 12 STR, 14 DEX, 12 WIL, claws (d6+d6) or bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Behave like felines: Curious, but dangerous if cornered. Skilled trackers and swimmers.
\item Hunt by sneaking on their prey.
\item \textbf{Critical Damage}: A human victim gets infected, becoming a lycanthrope of the same type after a couple of weeks.
\end{itemize}
\end{samepage}

\hypertarget{werewolf}{%
\section{Werewolf}\label{werewolf}}

4 HP, 12 STR, 14 DEX, 12 WIL, claws (d4+d4) or bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Semi-intelligent, ferocious hunters.
\item Sometimes hunts in packs with regular wolves.
\item \textbf{Critical Damage}: Victim becomes infected, transforming into a lycanthrope of the same type after a couple of weeks.
\end{itemize}
\end{samepage}

\hypertarget{wolf-dire}{%
\section{Wolf, Dire}\label{wolf-dire}}

5 HP, 14 STR, 12 DEX, bite (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, savage, semi-intelligent wolves. Dwell in caves, mountains, and forests.
\item Can be trained like dogs if captured young. Dire wolves are particularly difficult to train.
\item Sometimes are seem used as mounts by goblins.
\end{itemize}
\end{samepage}

\hypertarget{wolf-normal}{%
\section{Wolf, Normal}\label{wolf-normal}}

3 HP, 14 DEX, bite (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Dwell primarily in wildlands, but occasionally lair in caves.
\item Can be trained like dogs if captured young.
\item When found in packs of at least 4, treat WIL as 15 for morale checks.
\end{itemize}
\end{samepage}

\hypertarget{wood-troll}{%
\section{Wood Troll}\label{wood-troll}}

12 HP, 15 STR, 12 DEX, 7 WIL, claws and bite (d8+d8 \emph{Blast})

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item As an action, can recover lost HP.
\item \textbf{Critical Damage}: moss and twigs begin growing out of target’s wounds.
\end{itemize}
\end{samepage}

\hypertarget{wyvern}{%
\section{Wyvern}\label{wyvern}}

7 HP, 14 DEX, 11 WIL, bite (d8+d8) or poisonous sting (d6)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Winged, two-legged, dragon-like monsters with a long tail tipped with a venomous stinger. Dwell in any terrain, but favor cliffs and forests.
\item \textbf{Critical Damage}: The deadly poison inflicts d8 extra STR damage.
\end{itemize}
\end{samepage}

\hypertarget{yeti}{%
\section{Yeti}\label{yeti}}

4 HP, 14 STR, 12 WIL, claws (d6+d6) or paralyzing gaze (save)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Large, shaggy ape-like monsters with vicious claws and a taste for human flesh. Dwell in isolated, icy regions.
\item Camouflage in the ice and snow to surprise its victims.
\item Gaze: surprised victims must save WIL or be paralyzed.
\end{itemize}
\end{samepage}

\hypertarget{zombie}{%
\section{Zombie}\label{zombie}}

2 HP, 14 STR, 6 DEX, 3 WIL, weapon (d8)

\begin{samepage}
\begin{itemize}
\setlength\itemsep{-.5em}
\item Slow, mindless humanoid corpses, reanimated as guardians by  powerful clerics or wizards.
\item Attack anything that comes nearby.
\item Immune to mind altering effects.
\end{itemize}
\end{samepage}

\end{document}
